What's in a Domain? Anaylsis of URL Features

### 2022-10-08 paper/paper.tex
Expanding experiments to cover SPAM and DMOZ

The Uniform Resource Locator (URL) is a ubiquitous element the digital world. 
We use them to advertise and locate businesses, to retrieve news and media and 
to interact with other people across a broad of social applications.
Processing URLs is key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
In applications like malicious wesite detection the URL needs to be processed with
in a rapid and efficient manner to provide utility\cite{Ma2009b}.
Common use cases for URL centered analysis include security analysis of potenital phishing 
attacks\cite{Basnet2014,Mamun2016,Vazhayil2018}, and identification of pages 
that host malware or viruses \cite{Canali2011,Mamun2016}.
There are also applications to online advertising
including anticipation of conversions \cite{Qiu2020}, contextual analysis of
the content\cite{Kan2005,Baykan2009,Meshkizadeh2010} or language\cite{Baykan2013} of webpages.

@@

In this work we describe an ontology of URL fetaures that relates to underlying origin
and purpose. We then present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features.

@@

\cite{Ayes2019}
\cite{Li2020}
\cite{Basnet2012}
\cite{Basnet2014}
\cite{Tupsamudre2019}
\cite{Baykan2009}
\cite{Kan2005}
\cite{Hernandez2012}
\cite{Chung2010}
\cite{Qiu2020}
\cite{Xu2021}
\cite{Rhea2022}
\cite{Ma2009}
\cite{Verma2017}
\cite{Sountharrajan2020}
\cite{Murthy2022}

@@

Common approaches to feature engineering on URLs include string patterns and regular expressions to
identify key sequences and lexical properties \cite{Kan2004,Mamun2016}, 
creation of lookup tables or likelihood scores based on key sequences
\cite{Meshkizadeh2010}, n-gram or bag of words models \cite{Baykan2009}, 
the generation of task specific embedding vectors\cite{Le2018,Qiu2020} and the
usage of domain name servers or registrars for ancillary information about the domain registration \cite{Canali2011}. 
Some authors have 
emphasized that URLs are sequences of characters and require feature extraction methods that respect 
this sequential nature\cite{Le2018,Vazhayil2018}. They have naturally turned to developing neural network approaches
based on convolutional or recurrent layers for learning these sequential structures.
However, the ordering of key features within URLs is predominantly fixed as shown in Figure \ref{fig:url_structure}.
The core elements of URL structure occur at fixed positions within the sequence, in contrast to human language
sequences which are built from flexible grammars that allow variable positioning of most key elements.

@@

The features generated by the url2feature package are summairized in Table \ref{tab:features}

\begin{table*}
\caption{URL Features}
\label{tab:features}
\begin{tabular}{|l|l|l|l|}
\toprule
Group         &Feature              &Type        &Definition  \\
\midrule
Protocol      &protocol\_name         &Category    &The exact internet protocol name    \\
              &protocol\_type         &Category    &Internet protocol categorised by its purpose    \\
              &protocol\_exists       &Boolean     &Flag indicating presence or absence of internet protocol   \\
\midrule
Host          &host\_is\_ip           &Boolean     &Flag indicating if the host is an IP address    \\
              &host\_has\_port        &Boolean     &Flag indicating if a port number is explicitly specified    \\
              &domain\_sections       &Numeric     &Count of period separated domain sections    \\
              &domain\_reg\_year      &Numeric     &Year in which the domain was initially registered    \\
              &subdomain\_type        &Category    &Categorisation of the purpose of the subdomain \\
              &subdomain\_freq        &Numeric     &Frequency of the subdomain in internet traffics    \\
              &tld\_type              &Category    &Top Level Domain extension categorised by purpose    \\
              &tld\_freq              &Boolean     &Top Level Domain extension frequency in internet traffic   \\
\midrule
Path          &path\_depth            &Numeric     &The depth of the directory path between host and file   \\
              &path\_words            &Numeric     &Count of distinct words in the path between host and file   \\
              &path\_has\_date         &Boolean     &Flag indicating if a date is detected in the path  \\
              &path\_is\_home          &Boolean     &Flag indicating if the path starts with a home directory    \\
\midrule 
File          &file\_ext              &Category    &The exact file extension e.g. "exe" or "php"    \\
              &file\_type             &Category    &File extension categorised by function \& purpose    \\
              &file\_ext\_exists      &Boolean     &Flag indicating presence of a file extension for the URL target  \\
\midrule
Params        &params\_len            &Numeric     &The character length of the parameter string    \\
              &params\_count          &Numeric     &The count of the key value pairs in the param string    \\
              &params\_has\_url       &Boolean     &Flag indicating presence of a raw URL within the parameter string   \\
              &params\_enc\_url       &Boolean     &Flag indicating presence of an encoded URL within the parameter string   \\
              &params\_enc\_char      &Boolean     &Flag indicating presence of an encoded character within the parameter string   \\
              &params\_frag\_len      &Numeric     &Length of a hash delineated fragement at the end of the parameter string   \\
\bottomrule
\end{tabular}
\end{table*}

These URL feature functions are all available in the open source python package \emph{url2features}
available at https://pypi.org/project/url2features with links to usage documentation and source
code.

@@

\begin{equation}
\label{campaign_baseline}
{}^c\mathcal{A}_p =  \frac{ \sum_{i \in I_p} \mathbf{1}^c(i) A_i }{ \sum_{i \in I_p} \mathbf{1}^c(i) }
\end{equation}

@@

\begin{table}
\caption{Machine Learning Results}
\label{tab:results}
\begin{tabular}{|l|r|r|r|r|}
\toprule
Dataset              &NB     &LR    &XT    &LGBM     \\
\midrule
Malware              &-      &-     &-     &-        \\
Phishing             &-      &-     &-     &-        \\
WebKb 4Uni           &-      &-     &-     &-        \\
Syskill \& Webert    &-      &-     &-     &-        \\
\bottomrule
\end{tabular}
\end{table}

>>>>

The Uniform Resource Locator (URL) is a ubiquitous element the digital world. 
We use them to advertise and locate businesses, to retrieve news and media and 
to interact with other people across a broad of social applications.
Processing URLs is key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
In applications like malicious wesite detection the URL needs to be processed with
in a rapid and efficient manner to provide utility\cite{Ma2009b}.
Common use cases for URL centered analysis include security analysis of potenital phishing 
attacks\cite{Basnet2012,Basnet2014,Mamun2016,Vazhayil2018,Tupsamudre2019}, and identification of pages 
that host malware or viruses \cite{Canali2011,Mamun2016}.
There are also applications to online advertising
including anticipation of conversions \cite{Qiu2020}, contextual analysis of
the content\cite{Kan2005,Baykan2009,Meshkizadeh2010} or language\cite{Baykan2013} of webpages.

@@

In this work we develop a feature engineering library that maintains an ontology of URL features 
that relate to underlying structure of the URL, with a method for visualising the importance of features
across the URL structure. These techniques can prove useful for analysing features for problems that
consist of moving targets and require broad sets of potential features and efficient feature selection \cite{Basnet2012}. 
We then present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features.

@@

\cite{Ayes2019}
\cite{Li2020}
\cite{Baykan2009}
\cite{Kan2005}
\cite{Hernandez2012}
\cite{Chung2010}
\cite{Qiu2020}
\cite{Xu2021}
\cite{Rhea2022}
\cite{Ma2009}
\cite{Verma2017}
\cite{Sountharrajan2020}
\cite{Murthy2022}

@@

Common approaches to feature engineering on URLs include string patterns and regular expressions to
identify key sequences and lexical properties \cite{Kan2004,Mamun2016,Tupsamudre2019}, 
creation of lookup tables or likelihood scores based on key sequences
\cite{Meshkizadeh2010}, n-gram or bag of words models \cite{Baykan2009}, 
the generation of task specific embedding vectors\cite{Le2018,Qiu2020} and the
usage of domain name servers or registrars for ancillary information about the domain registration \cite{Canali2011}. 
Some authors have 
emphasized that URLs are sequences of characters and require feature extraction methods that respect 
this sequential nature\cite{Le2018,Vazhayil2018}. They have naturally turned to developing neural network approaches
based on convolutional or recurrent layers for learning these sequential structures.
However, the ordering of key features within URLs is predominantly fixed as shown in Figure \ref{fig:url_structure}.
The core elements of URL structure occur at fixed positions within the sequence, in contrast to human language
sequences which are built from flexible grammars that allow variable positioning of most key elements. Furthermore,
there are psychological and social elements to URL construction (such as the use of common brand names in subdomains
for phishing attack URLS \cite{Tupsamudre2019}), which emphasize the utility of feature engineering techniques that
respect the inherent structure of URLs.

@@

We generate URL features in groups that focus on distinct regions across the URL. We create a group
of global features that contain general string properties of the entire URL, then create regions
specific features as outlined in Table \ref{tab:features}.

\begin{table*}
\caption{URL Features}
\label{tab:features}
\begin{tabular}{|l|l|l|l|}
\toprule
Group         &Feature              &Type        &Definition  \\
\midrule
Protocol      &protocol\_name         &Category    &The exact internet protocol name    \\
              &protocol\_type         &Category    &Internet protocol categorised by its purpose    \\
              &protocol\_exists       &Boolean     &Flag indicating presence or absence of internet protocol   \\
\midrule
Host          &host\_is\_ip           &Boolean     &Flag indicating if the host is an IP address    \\
              &host\_has\_port        &Boolean     &Flag indicating if a port number is explicitly specified    \\
              &domain\_sections       &Numeric     &Count of period separated domain sections    \\
              &domain\_reg\_year      &Numeric     &Year in which the domain was initially registered    \\
              &subdomain\_type        &Category    &Categorisation of the purpose of the subdomain \\
              &subdomain\_freq        &Numeric     &Frequency of the subdomain in internet traffics    \\
              &tld\_type              &Category    &Top Level Domain extension categorised by purpose    \\
              &tld\_freq              &Boolean     &Top Level Domain extension frequency in internet traffic   \\
\midrule
Path          &path\_depth            &Numeric     &The depth of the directory path between host and file   \\
              &path\_words            &Numeric     &Count of distinct words in the path between host and file   \\
              &path\_wd\_len          &Numeric     &The mean length of the individual words in the path  \\
              &path\_has\_date        &Boolean     &Flag indicating if a date is detected in the path  \\
              &path\_is\_home         &Boolean     &Flag indicating if the path starts with a home directory    \\
\midrule 
File          &file\_len              &Numeric     &The character length of the target file    \\
              &file\_wds              &Numeric     &The count of words longer than 2 chars in the file name   \\
              &file\_wd\_len          &Numeric     &The mean length of the individual words in the file name  \\
              &file\_ext              &Category    &The exact file extension e.g. "exe" or "php"    \\
              &file\_type             &Category    &File extension categorised by function \& purpose (e.g. Static or dynamic)   \\
              &file\_ext\_exists      &Boolean     &Flag indicating presence of a file extension for the URL target  \\
\midrule
Params        &params\_len            &Numeric     &The character length of the parameter string    \\
              &params\_count          &Numeric     &The count of the key value pairs in the param string    \\
              &params\_has\_url       &Boolean     &Flag indicating presence of a raw URL within the parameter string   \\
              &params\_enc\_url       &Boolean     &Flag indicating presence of an encoded URL within the parameter string   \\
              &params\_enc\_char      &Boolean     &Flag indicating presence of an encoded character within the parameter string   \\
              &params\_frag\_len      &Numeric     &Length of a hash delineated fragement at the end of the parameter string   \\
\bottomrule
\end{tabular}
\end{table*}

These URL feature functions are available in our open source implmentation and python package \emph{url2features}. 
The package is designed such that it can process large files in chunks and will only add the specific sub-groups of
features a user requests.

@@

We use the SHAP\cite{Lundberg2017} package to calculate feature importance using a small holdout test set 
of approximately 50 smaples per experiment. We aggregate the shapley values for these sample points to calculate the mean
absolute contribution of each feature.

For our URL feature analysis, we then group these individual feature importance values by the URL segment
that belong to. We sum the feature importance for all features from that segment to allow us to plot the
segment contributions to model performance. These plots allow structural comparison features across the 
four problems in this study.

@@

A set of standard machine learning pipelines are applied to each of the four problems. Each approach is applied
using default parameters, without any fine tuning. We use feature preprocessing modules appropriate for each class
of model. All details are available in the source repository for the experiments. The performance of the models
is tabulated in Table \ref{tab:results}, where we show Balanced Accuracy of each model on the holdout data, as this
metric is appropriate and comparable across all four problems.

\begin{table}
\caption{Machine Learning Results}
\label{tab:results}
\begin{tabular}{|l|r|r|r|r|}
\toprule
Dataset              &NB     &LR    &XT    &LGBM     \\
\midrule
Malware              &0.65   &0.93  &0.99  &1.00     \\
Phishing             &0.89   &0.98  &0.99  &1.00     \\
WebKb 4Uni           &0.30   &0.36  &0.58  &0.49     \\
Syskill \& Webert    &0.39   &0.35  &0.37  &0.33     \\
\bottomrule
\end{tabular}
\end{table}

We take the best performing model for each problem and generate the SHAP values for the records in the test data.
These records are then used to generate the URL Feature Importance Plot shown in Figure \ref{fig:importance}.

\begin{figure*}
\centering
\includegraphics[scale=0.8]{images/URL_importance.png}
\caption{Structural Depiction of URL Feature Importance}
\label{fig:importance}
\end{figure*}
